MCC Predictions

### 2018-05-18 stats/ex3_comparisons.Rmd
continuing model comparisons

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
```

@@

```{r model predictions}
predictions <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"),
  Model = c("Full model","Full model","Full model","Full model","Full model","Full model","Full model","Full model","No prior","No prior","No prior","No prior","No prior","No prior","No prior","No prior", "Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only"),
  corr = c(0.9999672744165804,0.03460862921907267,0.9083230583454356,0.5090645649737232,0.9930727540990627,0.20045767964927574,0.8603373265403738,0.594975610001643,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.967,0.033,0.642,0.358,0.833,0.167,0.592,0.408))

@@

```{r joining files}

@@

```{r rsaUtils}
library(rwebppl)

rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

```{r rsa model2}
rsaModel2 <- '
var pragmaticListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
var S1 = speaker(obj, lexiconName, priorProbs);
observe(S1, utterance)
return obj.shape == "circle" ? 1 : 0
}})
}

@@

```{r bda model rsa noise}
bdaModel.rsa.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior noise}
bdaModel.prior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model rsa regularizedPrior}
bdaModel.rsa.regularizedPrior <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior regularizedPrior}
bdaModel.prior.regularizedPrior <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

@@

```{r bda model rsa regularizedPrior noise}
bdaModel.rsa.regularizedPrior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior regularizedPrior noise}
bdaModel.prior.regularizedPrior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r model setup}

@@

```{r predictions rsa noise}
bda <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions prior noise}
bda1 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions rsa regularized prior}
bda2 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.regularizedPrior, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions prior regularized prior}
bda3 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.regularizedPrior, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions rsa noise regularized prior}
bda4 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.regularizedPrior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

noisy.regPrior.rsa.model <- posterior.predictive4 %>%
  mutate(Model="Noisy Regularized RSA Model",
         Adjustment = "Noise & Regularized Prior",
         Type = "RSA")%>%
  group_by(Model,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))

```{r predictions prior noise regularized prior}
bda5 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.regularizedPrior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r}
## joining noisy model predictions with other predictions
no.adj.models <- pd %>%
  mutate(Type = ifelse(Model == "Full model" | Model == "No prior","RSA",ifelse(Model == "Data","Data","Prior only")),
         Adjustment = "None")

@@

ggplot(p.com, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Model, labeller = label_wrap_gen(width=5))+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)

@@

plot.model <- p.com %>%
  filter(Model != "Data", Model != "No prior" )

@@

```{r model rsa loglikelihood}
bdaModel.rsa.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

  //display(JSON.stringify(modelPredictions.score(d.correct_inf)))
  //display(JSON.stringify(conditionSpecificData))

@@

 return sum(loglike)
}
'
```

```{r rsa loglikelihood}
llh <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 100, verbose = T)
)

@@

>>>>

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
library(rwebppl)
```

@@

```{r model predictions, include = FALSE}
predictions <- data.frame(
  Experiment = c("Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty","Preference","Preference","Preference","Preference","Novelty","Novelty","Novelty","Novelty"),
  Change = c("Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker","Same speaker","Same speaker","Different speaker","Different speaker"),
  Alignment = c("Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent","Congruent","Incongruent"),
  Model = c("Full model","Full model","Full model","Full model","Full model","Full model","Full model","Full model","No prior","No prior","No prior","No prior","No prior","No prior","No prior","No prior", "Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only","Prior only"),
  corr = c(0.9999672744165804,0.03460862921907267,0.9083230583454356,0.5090645649737232,0.9930727540990627,0.20045767964927574,0.8603373265403738,0.594975610001643,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.742,0.967,0.033,0.642,0.358,0.833,0.167,0.592,0.408))

@@

```{r joining files, include = FALSE}

@@

```{r rsaUtils, include = FALSE}
rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

@@

```{r rsa model2, include = FALSE}
rsaModel2 <- '
var pragmaticListener = function(utterance, priorProbs){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample( Categorical({vs: all_objects, ps: priorProbs}));
var S1 = speaker(obj, lexiconName, priorProbs);
observe(S1, utterance)
return obj.shape == "circle" ? 1 : 0
}})
}

@@

```{r bda model rsa noise, include = FALSE}
bdaModel.rsa.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior noise, include = FALSE}
bdaModel.prior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model rsa regularizedPrior, include = FALSE}
bdaModel.rsa.regularizedPrior <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior regularizedPrior, include = FALSE}
bdaModel.prior.regularizedPrior <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

   map(function(d){ 
      observe(modelPredictions, d.correct_inf)
    }, conditionSpecificData)

@@

```{r bda model rsa regularizedPrior noise, include = FALSE}
bdaModel.rsa.regularizedPrior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r bda model prior regularizedPrior noise, include = FALSE}
bdaModel.prior.regularizedPrior.noise <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

```{r model setup, include = FALSE}

@@

```{r predictions rsa noise, include = FALSE}
bda <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions prior noise, include = FALSE}
bda1 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions rsa regularized prior, include = FALSE}
bda2 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.regularizedPrior, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions prior regularized prior, include = FALSE}
bda3 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.regularizedPrior, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r predictions rsa noise regularized prior, include = FALSE}
bda4 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.rsa.regularizedPrior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

noisy.regPrior.rsa.model <- posterior.predictive4 %>%
  mutate(Model="Noisy Regularized RSA Model",
         Adjustment = "Noise & Regularized Prior",
         Type = "RSA")%>%
  group_by(Model,Adjustment,Type,Experiment,Change,Alignment) %>%
  summarise(mean = estimate_mode(value), ci_lower = hdi_lower(value), ci_upper = hdi_upper(value))
```

```{r predictions prior noise regularized prior, include = FALSE}
bda5 <- webppl(
  program_code = paste(rsaUtils, rsaModel2, bdaModel.prior.regularizedPrior.noise, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "MCMC", samples = 6000, burn = 2000, verbose = T)
)

@@

```{r}
## joining noisy model predictions with other predictions
no.adj.models <- pd %>%
  mutate(Type = ifelse(Model == "Full model","RSA",ifelse(Model == "Data","Data",ifelse(Model == "No prior","No Prior","Prior only"))),
         Adjustment = "None")

@@

p.com$Adjustment <- ordered(p.com$Adjustment, levels = c("None","Noise","Regularized Prior","Noise & Regularized Prior"))

ggplot(p.com, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Adjustment*Type, labeller = label_wrap_gen(width=5))+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)

@@

plot.model <- p.com %>%
  filter(Model != "Data", Model != "No prior")

plot.model$Adjustment <- ordered(plot.model$Adjustment, levels = c("None","Noise","Regularized Prior","Noise & Regularized Prior"))

@@

```{r model rsa loglikelihood}
rsa.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

   var modelPredictions = pragmaticListener({label: "dax", point: 2 }, conditionSpecificPriors[0].priorProbs)

 return sum(conditionOutput)
}
'
```

```{r model prior loglikelihood}
prior.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

  return priorOnlyPredictions.score(d.correct_inf)
    }, conditionSpecificData))

 return sum(conditionOutput)
}
'
```

```{r model rsa noise loglikelihood}
rsa.noise.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

  return noisyModelPredictions.score(d.correct_inf)
    }, conditionSpecificData))

 return sum(conditionOutput)
}
'
```

```{r model prior noise loglikelihood}
prior.noise.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

  return noisyModelPredictions.score(d.correct_inf)
    }, conditionSpecificData))

 return sum(conditionOutput)
}
'
```

```{r model rsa regularized prior loglikelihood}
rsa.regPrior.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

@@

@@

 return sum(conditionOutput)
}
'
```

```{r model prior regularized prior loglikelihood}
prior.regPrior.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

 return sum(conditionOutput)
}
'
```

```{r model rsa regularizedPrior noise loglikelihood}
rsa.regularizedPrior.noise.loglikelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

  return noisyModelPredictions.score(d.correct_inf)
    }, conditionSpecificData))

 return sum(conditionOutput)
}
'
```

```{r model prior regularizedPrior noise loglikelihood}
prior.regularizedPrior.noise.likelihood <-'
var allConditions = dataFromR.allConditions
var allData = dataFromR.allData
var allPriorProbs = dataFromR.allPriorProbs

  return noisyModelPredictions.score(d.correct_inf)
    }, conditionSpecificData))

 return sum(conditionOutput)
}
'
```

```{r model loglikelihoods}

llh.rsa <- webppl(
  program_code = paste(rsaUtils, rsaModel2, rsa.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 5000, chains = 3, cores = 3)
)%>%
  mutate(Model = "RSA")

llh.prior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, prior.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 5000, chains = 3, cores = 3)
)%>%
  mutate(Model = "Prior only")

llh.rsa.noise <- webppl(
  program_code = paste(rsaUtils, rsaModel2, rsa.noise.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
)%>%
  mutate(Model = "RSA Noisy")

llh.prior.noise <- webppl(
  program_code = paste(rsaUtils, rsaModel2, prior.noise.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
)%>%
  mutate(Model = "Prior only Noisy")

llh.prior.regPrior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, prior.regPrior.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
)%>%
  mutate(Model = "Prior only Regularized")

llh.rsa.regPrior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, rsa.regPrior.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
)%>%
  mutate(Model = "RSA Regularized")

llh.rsa.noise.regPrior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, rsa.regularizedPrior.noise.loglikelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
)%>%
  mutate(Model = "RSA Noisy Regularized")

llh.prior.noise.regPrior <- webppl(
  program_code = paste(rsaUtils, rsaModel2, prior.regularizedPrior.noise.likelihood, sep='\n'),
  data = list(allData = data, allConditions = allConditions, allPriorProbs = allPriorProbs), 
  data_var = "dataFromR",
  model_var = "model",
  inference_opts = list(method = "forward", samples = 1e+5, chains = 3, cores = 3)
) %>%
  mutate(Model = "Prior only Noisy Regularized")

@@

```{r}

likelihood_overview <- bind_rows(
  llh.rsa,
  llh.prior,
  llh.rsa.noise,
  llh.prior.noise,
  llh.prior.regPrior,
  llh.rsa.regPrior,
  llh.rsa.noise.regPrior,
  llh.prior.noise.regPrior)

ggplot(likelihood_overview, aes(x = value))+
  geom_histogram()+
  xlim(-700,-500)+
  facet_wrap(~Model, nrow = 2)

```{r logsum summary}
library(matrixStats)

llh.plot <- likelihood_overview %>%
  group_by(Model)%>%
  summarize(logSumExp = logSumExp(value))%>%
  mutate(Model = reorder(Model,logSumExp))

ggplot(llh.plot, aes(x = Model, y = logSumExp, fill = Model))+
  geom_bar(stat="identity", color = "black")+
  ggtitle("logsumexp per model")+
  theme_few()+ 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))

save.image(file="model.comparisons.RData")
```
